---
title: "Proxy Carbon Data Cleaning and Figure 2,3,4 generation"
author: "S Sayre and A Barron with assistance from Chhiring Lama"
date: "6/24/20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,warning = FALSE, message = FALSE}
#loading the packages ----
require(dplyr)
require(tidyverse)
require(ggthemes)
require(lubridate)
require(readxl)
require(data.table)
require(gridExtra)
```

Note: Dillon gym analysis is a LCC sensitivity on first cost so can't be used for most analysis (that requires a known total cost) and should be dropped.

Loading Data and tidying for fig 2
```{r, warning = FALSE, message = FALSE}
SmithProxyCarbon_LCC <- read_excel("SI_Figure_2_Smith_ProxyCarbon_LCC.xlsx", sheet = 1) 
names(SmithProxyCarbon_LCC)[1]<- "sector" 

SmithProxyCarbon_LCC <- SmithProxyCarbon_LCC %>%
  gather(key = "key", value = "value", -sector) #tidying the data for grouped bar charts
```

Fig 2:
```{r make fig2}
tiff("fig2.tiff",res = 300, unit = "in",height = 6, width =8)
collist <- c("#E1BE6A","#63ABA4","#AF7171")

fig2 <- ggplot(SmithProxyCarbon_LCC, aes(x = key, y = value, fill = factor(sector, levels=c("Carbon", "Energy", "Capital"))))+
  geom_bar(width = 0.6, position = "stack", stat = "identity")+ 
  scale_fill_manual(name = "", values = collist)+
  scale_y_continuous(name = "Net Present Value Cost ('000 $2017)",limits = c(0,400),expand = c(0, 0), breaks = c(0,50,100,150,200,250,300,350, 400), labels = c("$0","$50","$100","$150","$200","$250","$300","$350", "$400"))+
  scale_x_discrete(name = "",breaks = c("Baseline", "Insulate Attic Plus Basement","Plus Above Grade Walls", "Plus Air Sealing", "Plus New Windows"), labels = c("Baseline", "Insulate Attic\n Plus Basement","Plus Above\n Grade Walls", "Plus Air\n Sealing", "Plus New\n Windows"))+
  theme_clean()+
  theme(plot.title = element_text(size = 16),
        axis.title.y = element_text(size = 16, margin = margin(t = 0, r = 20, b = 0, l = 0)), 
        axis.text = element_text(size = 12),
        legend.text = element_text(size = 12), 
        legend.position = c(0.5, 0.815),
        legend.margin = margin(t = -10, r = 5, b = 5, l = 5))
print(fig2)
dev.off()
```

Loading Data for fig 3a
```{r, warning = FALSE, message = FALSE}
PrincetonLCC <- read_excel("Princeton_Master_R_01_22_19.xls")
#As of 10/25/18 should be 55 rows for 71 vars
#removing springdale clubhouse which is pre-ICP
PrincetonLCC <-PrincetonLCC %>% filter(project_name != "springdale_clubhouse")
#dropping dillon gym due to non-standard analysis structure (sensivity)
#dropping childcare windows and acee (no initital cost)
#dropping two lighting analysis (LED vs alt fixtures) - no carbon price use
#dropping chilled water LCC - No carbon price used
PrincetonLCC_all<-PrincetonLCC
PrincetonLCC <- PrincetonLCC %>% filter(project_name != "dillon_gymnasium") %>%
filter(project_name != "university_childcare")  %>%
  filter(project_name != "acee") %>%
  filter(study_category != "chilled_water") %>%
  filter(alternates_options != "alt_fixtures") %>%
  filter(alternates_options != "dd_led") %>%
  filter(project_name != "insulation_roof_existing" & !is.na(pv_ctax_energy)) 
```


General cleanup
```{r, warning = FALSE, message = FALSE}
PrincetonLCC$total_lcc <- as.numeric(PrincetonLCC$total_lcc)
PrincetonLCC <- PrincetonLCC %>% mutate(project_size = replace(project_size, project_size == c("7650 SQ-FT"), 7650))
#temp clean immediate dates from fields
PrincetonLCC <- PrincetonLCC %>% mutate(base_date = replace(base_date, base_date == c("immediate"), NA)) %>% mutate(service_date = replace(service_date, service_date == c("immediate"), NA))
PrincetonLCC$base_date <- as.numeric(PrincetonLCC$base_date)
PrincetonLCC$service_date <- as.numeric(PrincetonLCC$service_date)
PrincetonLCC$base_date <- as.Date(PrincetonLCC$base_date, origin = "1899-12-30")
PrincetonLCC$service_date <- as.Date(PrincetonLCC$service_date, origin = "1899-12-30")
PrincetonLCC$base_date <- as.POSIXct(PrincetonLCC$base_date)
PrincetonLCC$service_date <- as.POSIXct(PrincetonLCC$service_date)
PrincetonLCC <- PrincetonLCC %>% mutate(date = replace(date, date == "2000-01-02 UTC", "2008-10-01 UTC"))
```

Fixing bad spreadsheet refs
```{r, warning = FALSE, message = FALSE}
#Photovoltaic HibbenMagie had bad ref to PV for Ctax
PrincetonLCC <- PrincetonLCC %>% mutate(pv_ctax_energy = replace(pv_ctax_energy, which(pv_ctax_energy == 711522.65), 102175.22))
PrincetonLCC <- PrincetonLCC %>% mutate(pv_ctax_energy = replace(pv_ctax_energy, which(pv_ctax_energy == 1450063.84), 153103.39))
#Building Envelope HibbenMagie had bad cell ref to PV for Ctax for alt 2 (did not go to alt c tax G vs J)
PrincetonLCC <- PrincetonLCC %>% mutate(pv_ctax_energy = replace(pv_ctax_energy, which(pv_ctax_energy == 3287957.87), 641781))
PrincetonLCC <- PrincetonLCC %>% mutate(pv_ctax_energy = replace(pv_ctax_energy, which(pv_ctax_energy == 3419399.16), 667004))
#Building Envelope HibbenMagie had bad cell ref to PV for Ctax for alt 1 - cited ~total for alt 2)
PrincetonLCC <- PrincetonLCC %>% mutate(pv_ctax_energy = replace(pv_ctax_energy, which(pv_ctax_energy == 3271528.42), 795340))
PrincetonLCC <- PrincetonLCC %>% mutate(pv_ctax_energy = replace(pv_ctax_energy, which(pv_ctax_energy == 3402312.91), 827135))
```

Cleaning up missing annual_energy_cost and pv_energy+cost
```{r, warning = FALSE, message = FALSE}
#Given that discount rate and energy escalation are equal in this dataset, approximating missing values by dividing by study period
PrincetonLCC <- PrincetonLCC %>% mutate(annual_energy_cost_interp = ifelse(is.na(annual_energy_cost),pv_energy_cost/study_period ,annual_energy_cost))%>%
mutate(pv_energy_cost_interp = ifelse(is.na(pv_energy_cost),annual_energy_cost*study_period ,pv_energy_cost))
```

Generating reference case values
```{r, warning = FALSE, message = FALSE}
PrincetonLCC_bl <- PrincetonLCC %>%
  filter(baseline == 1) %>%
  select(Case,initial_cost,pv_Investment_Replacement_LCC,annual_energy_cost_interp,pv_energy_cost_interp,pv_ctax_energy,pv_energy_ctax_lcc,pv_total_lcc,pv_m_r_lcc) %>%
  rename(initial_cost_bl = initial_cost)%>%
  rename(annual_energy_cost_interp_bl = annual_energy_cost_interp)%>%
  rename(pv_Investment_Replacement_LCC_bl = pv_Investment_Replacement_LCC)%>%
  rename(pv_energy_cost_interp_bl = pv_energy_cost_interp)%>%
  rename(pv_ctax_energy_bl = pv_ctax_energy)%>%
  rename(pv_energy_ctax_lcc_bl = pv_energy_ctax_lcc)%>%
  rename(pv_total_lcc_bl = pv_total_lcc) %>%
  rename(pv_m_r_lcc_bl = pv_m_r_lcc)
#Adding baseline values back in to table
PrincetonLCC <- PrincetonLCC %>% left_join(PrincetonLCC_bl,by="Case")
```

#creating variables for Figure 3 and exporting table
```{r, warning = FALSE, message = FALSE}
#Pull out baseline cases and add relevant variables
princeton_data <- PrincetonLCC %>%
  filter(baseline == 0) %>%
#  mutate(percent.red.energy.use=1-annual_energy_cost_interp/annual_energy_cost_interp_bl) %>%
#  mutate(initial.over.annual.energy=initial_cost/annual_energy_cost_interp) %>%
#  mutate(icp.over.icp.plus.energy=pv_ctax_energy/pv_energy_ctax_lcc) %>%
#  mutate(initital.over.pv.energy.bl=initial_cost/pv_energy_cost_bl) %>%
  select(Case,project_name,study_category,alternates_options,study_period,alternates_options,initial_cost,pv_m_r_lcc, pv_m_r_lcc_bl,pv_energy_cost_interp, pv_energy_cost_interp_bl, pv_ctax_energy, pv_ctax_energy_bl, pv_energy_ctax_lcc, pv_energy_ctax_lcc_bl,pv_Investment_Replacement_LCC,pv_Investment_Replacement_LCC_bl, study_period, annual_energy_cost_interp, annual_energy_cost_interp_bl) %>%
  mutate(caseID = row_number())

#folder <- "/Users/abarron/Google Drive/"
#write.csv(PrincetonLCCf3, #file=file.path(folder,"PrincetonLCC_Fig3_clean020119.csv"))
```

```{r import and combine Princeton and Smith data, warning = FALSE, message = FALSE}
princeton_data <- princeton_data %>%
  mutate(project_cost = pv_Investment_Replacement_LCC - 
           pv_Investment_Replacement_LCC_bl + pv_m_r_lcc - pv_m_r_lcc_bl) %>%
  mutate(carbon_base = pv_energy_cost_interp_bl + pv_ctax_energy_bl,
         carbon_alt = pv_energy_cost_interp  + pv_ctax_energy,
         nocarbon_base = pv_energy_cost_interp_bl,
         nocarbon_alt = pv_energy_cost_interp)

breakeven <- princeton_data %>% 
  mutate(nocarbon_savings = pv_energy_cost_interp_bl - pv_energy_cost_interp,
         ctax_savings = pv_ctax_energy_bl - pv_ctax_energy,
         be_ratio = (project_cost - nocarbon_savings - ctax_savings)/ctax_savings)
           

graph_data <- princeton_data%>%
  select(caseID, alternates_options, study_period, project_cost, contains("carbon")) %>%
  gather(key = "type", value = "annual_cost", contains("carbon")) %>%
    separate(type, into = c("cost_type", "scen_type"), sep = "_") %>%
    spread(key = scen_type, value = "annual_cost") %>%
    mutate(invest_ratio = project_cost/base, 
           percent_reduction = 1 - alt/base,
           school = "Princeton")
#    filter(!caseID %in% c(1,2,11,12,14))

smith_data <- read_csv("smith_data.csv") %>% 
  gather(key = "alternates_options", value = "npv_cost", -cost_type) %>% 
  mutate(cost_category = case_when(cost_type == "utility" ~ "nocarbon",
                                   cost_type == "carbon" ~ "carbon_only",
                                   TRUE ~ "invest")) %>% 
  group_by(alternates_options, cost_category) %>% 
  summarize(npv_cost = sum(npv_cost)) %>% 
  spread(key = cost_category, value = "npv_cost") %>% 
  mutate(carbon = nocarbon + carbon_only) %>% 
  select(alternates_options, carbon, nocarbon, invest) 

baseline_data <- smith_data %>% 
  select(carbon_bl = carbon, nocarbon_bl = nocarbon, invest_bl = invest, baseline_case = alternates_options)

smith_compare_cases <- smith_data %>% 
  filter(alternates_options!="baseline" | alternates_options!="gasoline_cart") %>% 
  mutate(baseline_case = case_when(alternates_options == "basement" ~ "baseline",
                                   alternates_options == "insulation" ~ "basement",
                                   alternates_options == "sealing" ~ "insulation",
                                   alternates_options == "windows" ~ "sealing",
                                   alternates_options == "LiION_cart" ~ "gasoline_cart",
                                   alternates_options == "PbAcid_cart" ~ "gasoline_cart"),
         caseID = case_when(alternates_options == "basement" ~ 101,
                            alternates_options == "insulation" ~ 102,
                            alternates_options == "sealing" ~ 103,
                            alternates_options == "windows" ~ 104,
                            alternates_options == "LiION_cart" ~ 105,
                            alternates_options == "PbAcid_cart" ~ 106)) %>% 
  left_join(baseline_data) %>% 
  mutate(net_invest = invest - invest_bl,
         percent_carbon = 1-carbon/carbon_bl,
         percent_nocarbon = 1 - nocarbon/nocarbon_bl,
         ir_carbon = net_invest/carbon_bl,
         ir_nocarbon = net_invest/nocarbon_bl) %>% 
  select(caseID, alternates_options, baseline_case, contains("percent"), contains("ir")) %>% 
  gather(key = ratio_type, value = "ratio", contains("carbon")) %>% 
  separate(ratio_type, into = c("ratio_type", "cost_type"), sep = "_") %>% 
  spread(key = ratio_type, value = "ratio") %>% 
  rename(invest_ratio = ir, percent_reduction = percent) %>% 
  mutate(school = "Smith")

combined_data <- graph_data %>% 
  select(school, caseID, alternates_options, cost_type, invest_ratio, percent_reduction) %>% 
  bind_rows(smith_compare_cases) %>% 
  mutate(graph_type = recode(factor(cost_type, levels = c("nocarbon", "carbon")),
                               nocarbon = "without carbon price",
                               carbon = "with carbon price")) 
```

```{r make fig3a, warning = FALSE, message = FALSE, fig.height=5, fig.width=5}
fig3a <- ggplot(combined_data, aes(invest_ratio, percent_reduction)) +
  geom_point(aes(shape = graph_type, color = school), size = 3) +
  geom_line(aes(invest_ratio, percent_reduction, group = caseID, color = school)) +
  geom_line(aes(percent_reduction,percent_reduction),color = "black") + theme_bw(base_size = 16) + 
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
        legend.position = c(.2,.6), 
        legend.background = element_rect(fill = "white", size = .5, 
                                         linetype = "solid", color = "black"),
        legend.title = element_blank())+
  xlim(-.5,1.5)
  
fig3a_labeled <- fig3a +
  labs(x = "Ratio of investment to baseline lifetime energy cost",
       y = "Percent reduction in energy cost relative to baseline",
       title = "(a)") +
  scale_colour_manual(values = c ("#e77500","#002855")) +
  annotate("text", x = 1.12, y = .92, label = "break-even line")

#write_csv(graph_data,"fig3_graph_data.csv")
```

For figure 3b
```{r cornell data, warning = FALSE, message = FALSE,fig.height=5, fig.width=5}
cornell_data <- tribble(
 ~alt, ~invest, ~operating, ~offsets, ~methane,
 1,	47,	24,	0, 0,
 2,	50,	22,	0, 0,
 3,	62,	28,	0, 0,
 4,	55,	26,	0, 0,
 5,	42,	34,	0, 0,
 7,	31,	36,	2, 10,
 8,	36,	34,	2, 10,
 9,	28,	47,	4, 17,
 10, 34,	40,	3, 13) %>%
mutate(carbon = operating + offsets,
       carbon_methane = operating + offsets + methane) %>%
  select(alt,invest, no_carbon = operating, carbon, carbon_methane) %>% 
  gather(key = "type", value = "annual_cost", contains("carbon")) %>% 
  mutate(base_cost = case_when(type=='no_carbon' ~ 42,
                               type=="carbon" ~ 52,
                               TRUE ~ 85),
         invest_ratio = invest/base_cost,
         reduction = 1-annual_cost/base_cost,
         graph_type = recode(factor(type, levels = c("no_carbon", "carbon", "carbon_methane")),
                             no_carbon = "without carbon price",
                             carbon = "with carbon price (direct)",
                             carbon_methane = "with carbon price (direct+methane)"))


cornellplot <- 
ggplot(cornell_data, aes(invest_ratio, reduction)) +
  geom_point(aes(shape = graph_type), color = "red", size = 3) +
  geom_line(aes(invest_ratio, reduction, group = alt), colour = "red") +
  geom_line(aes(reduction,reduction),color = "black") + theme_bw(base_size = 16) + 
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
        legend.position = c(.75,.14), 
        legend.background = element_rect(fill = "white", size = .5, 
                                         linetype = "solid", color = "black"),
        legend.title = element_blank(),
        legend.margin = margin(t = -6, r = 3, b = 4, l = 3))

cornellplot <- cornellplot +
  labs(x = "Ratio of investment to baseline lifetime operating cost",
       y = "Percent reduction in operating cost relative to baseline",
       title = "(b)") +
  annotate("text", x = .9, y = .75, label = "break-even line")

```

Combining Figure 3a and 3b
```{r combine fig3, warning= FALSE, message=FALSE}
tiff("fig3.tiff", res = 300,unit = "in", height =7, width =15)
fig3 <- grid.arrange(fig3a_labeled, cornellplot, ncol = 2)
dev.off()
```

Loading data for Figure 4
```{r, warning= FALSE, message=FALSE}
proxy_prices <- read_excel("SI_Figure 4 Energy Price with ICP 6 23 20.xlsx", sheet = "Data")
proxy_prices_all<- proxy_prices
```

Cleaning up the table
```{r, warning= FALSE, message=FALSE}
proxy_prices <- proxy_prices[15:23,] #only selecting rows with refined data on price where the unit of the energy for all categories is MMBTU
proxy_prices <- proxy_prices %>%
  select_if(~ !any(is.na(.))) #delete all columns with no variable names as either the cols are empty or they have units (not required here)
names(proxy_prices) <- proxy_prices %>% slice(1)%>%unlist() 
proxy_prices <- proxy_prices[, !duplicated(colnames(proxy_prices))] %>%
  slice(-1) #removing the first column as it was the name of the variables in the raw data
oldnames <- names(proxy_prices)
newnames <- c("Category","Unit","MTCO2e/MMBTU","Energy_price", "Carbon_price", "Conversion_factors")
proxy_prices <-proxy_prices %>%rename_at(vars(all_of(oldnames)), ~ newnames)%>%#renaming the variables to use names that can be used in operations 
  gather(key = "Key", value = "Price", - c("Category","Unit","MTCO2e/MMBTU", "Conversion_factors"))%>% #tidying the data for the graph
  mutate(Unit ="$/MMBTU")
proxy_prices$Price <- as.numeric(proxy_prices$Price) #converting the values to a numeric value
```

```{r fig 4, message=FALSE, warning = FALSE}
tiff("fig4.tiff",res = 300, unit = "in",height = 6, width =8)

colors = c("#E1BE6A","#63ABA4")
fig4 <- ggplot(proxy_prices, aes(x = Category, y = Price, group = Key, fill = Key))+
  geom_bar(width = 0.6,position= "stack", stat = "identity") +
  scale_y_continuous(name = "$/MMBTU",limits = c(0,60),expand = c(0,0),breaks = c(0,10,20,30,40,50,60), labels = c("$0","10","$20","$30","$40","$50","$60"))+
  scale_x_discrete(name = "",breaks = c("Electricity         (Campus CHP)", "Electricity           (Northeast US)", "Electricity           (South East US)","Natural Gas (Commercial)", "Diesel","Gasoline", "Fuel Oil #2","Fuel Oil #6"), labels = c("Electrcity\n (Campus\n CHP)", "Electricity\n (NE US)", "Electricity\n (SE US)", "Natural Gas", "Diesel", "Gasoline","Fuel Oil\n #2", "Fuel Oil\n #6" ))+
  scale_fill_manual(name ="",breaks = c("Carbon_price", "Energy_price"), labels = c("Carbon Price","Energy Price"),values = colors)+
  theme_clean()+
  theme(axis.text = element_text(size =11),
        axis.title = element_text(size =16, margin = margin(t = 0, r = 10, b = 0, l = 0)), 
        legend.position = c(0.7,0.75),
        legend.margin = margin(t = -10, r = 5, b = 5, l = 5))
print(fig4)

dev.off()
```

